Delone lattice studies in \CIII, the space of three complex variables

### 2023-03-09 C3/C3.tex
C3 -- add the generating command to generate the graphics
updating to (I hope) final version

		The Delone (Selling) scalars, which are used in 
		unit cell reduction and in lattice type determination,
		are studied in \CIII, the space of 3 complex variables.
		The 3 complex coordinate planes are composed of the 6
		Delone scalars.

@@

	~\citeasnoun{andrews2019b} chose to 
	represent the Selling scalars in the space \SVI{},
	\Svec{} (defined in the order above), 
	as a way to create a metric space
	for the measurement of the distance between lattices. 
	They also consider the representation of this space as the
	space of three complex dimensions, \CIII{} or 
	{$\{c_1$}, {$c_2$}, {$c_3\}$}.

	In \CIII{}, in terms of the Selling scalars, 
	a vector is defined as \{(\si,\siv ), (\sii,\sv),(\siii,\svi)\}, 
	where the real and imaginary parts
	of each are the ``opposite'' scalars 
	according to the definition of~\citeasnoun{Delaunay1932} (see~\citeasnoun{andrews2019a}).
	As a mnemonic device, 
	note that the complex components involve ($\alpha,a$), ($\beta, b$), and ($\gamma,c$).
	Additionally, each complex term uses all 
	four 3-space vectors; for example, $c_1$ is ($b \cdot c$, $a \cdot d$).

	~\citeasnoun{andrews2019b} considered the matrix representations 
	of the reflections in \CIII{} (and in \SVI{}). 
	This paper describes the boundary transformations 
	at the edges of the fundamental	unit of \CIII{}. 
	In \SVI{}, the fundamental unit is the all negative orthant, 
	which contains only and	all of the reduced cells. 
	In \SVI{} and \CIII{}, 
	the boundaries located where any \SVI{} scalar 
	(or correspondingly in\CIII{}, the real or imaginary part) 
	equals to zero. The rationale for this work was 
	that it might lend insights into the topology of
	the space of lattices.

@@

	Next we define the operators in \CIII{} t
	hat will be used in the matrix descriptions of 
	the transformations at the boundaries of the fundamental unit.\\
	See Table ~\ref{tab:operators}.

@@

	\section{Matrices of boundary transformations}

@@

	~\citeasnoun{Delaunay1932} does not 
	consider the boundary transformations 
	in detail. 
	However, he uses them to 
	define the process of Selling reduction. 
	For example in \SVI{}, 
	he lists the following as one of the possible results 
	for a transformation on \si{}:
	\SvecA{}.
	The third boundary transform for $s_1$
	implements this operation and interchanges the
	real part of \ciii{} and the imaginary part of \cii{}:

@@

	Considered in \CIII, Delone's alternate transformation 
	for the \si{} boundary would
	exchange the real of \cii{} 
	with the imaginary part of \ciii{}. That is the fourth
	matrix in the list for \si{}. 
	The other two transformations for \si{} can 
	be generated from the two we have 
	just displayed by the "exchange operation"~\cite{andrews2019b} 
	applied to the second and third \CIII{} coordinates. Delone 
	did not describe the latter two transformation, 
	perhaps because even a single transformation 
	was adequate to implement reduction. He already listed two.

@@

	The 2-dimensional nature of the three coordinates of \CIII{} 
	suggests their use for graphical display.

	As an example, we use Phospholipase A2 (retrieved from 
	the Protein Data Bank \cite{Bernstein1977}), which has had several similar
	or identical structures determined \cite{LeTrong2007}. \citeasnoun{andrews2019b}
	found additional cases (see Table \ref{PLA2})

@@

Below, Figure \ref{fig1} shows the unit cells as 
reported (the centering of lattices has not been removed).
The following figures show various transformations and
embellishments of the reported cells.

@@

\begin{figure}
	\includegraphics[width=1.0\textwidth]{2_06}
	\caption{The unit cells Niggli reduced. The similarity 
	of the 3 projections is indicative of the exact or nearly
exact rhombohedral symmetry.}
	\label{fig2}
\end{figure}
\begin{figure}
	\includegraphics[width=1.0\textwidth]{3_03}
	\caption{The unit cells Delone reduced. }
	\label{fig3}
\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{4_13}
	\caption{The unit cells, Niggli and 5 copies were perturbed 2\%
	orthogonally to \SVI{} vector. }
	\label{fig4}
\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{5_21}
	\caption{The unit cells Niggli, and 5 copies were perturbed 10\%
		orthogonally to \SVI{} vector. }
	\label{fig5}
\end{figure}

@@

The transformation matrices shown above demonstrate the 
considerable regularity of Selling reduction, as used by 
Delaunay (1932), in comparison to Niggli 
reduction \cite{Niggli1928}. While all the boundaries of the non-positive 
orthant of \SVI{} are essentially the same (and 
related by reflections), the boundaries formed in 
Niggli reduction are of multiple types, and the fundamental unit 
of the space representing Niggli reduced cells (\GVI{}, see \citeasnoun{Andrews2014}) is non-convex.

The matrices also display one of the unique properties of \CIII{}, 
which makes it a useful conceptual representation. 
For the scalar being transformed, it and its opposite scalar
create a unique row in each transformation matrix. These rows contain only
the minus operator and zeros, highlighting the 
unique relationship of that pair of scalars.

Several aspects of \CIII{} are evident from inspecting the matrices. 
First, each boundary has four possible 
transformations that can be applied. Since each of 
the transformations at boundaries are self-inverse, they are the 
same transformations that would be used in the process of 
cell (lattice) reduction. \citeasnoun{Delaunay1932} and \citeasnoun{Delone1975} give only two choices, presumably for simplicity, 
omitting transformations that use the "exchange" operator 
(see \citeasnoun{andrews2019b}).

@@

	The $C^{++}$ ~code for \CIII{} is available in github.com, in
	\url{https://github.com/duck10/LatticeRepLib.git}.

@@

	\bibliography{reduced}

@@

>>>>

		The Delone (Selling) scalars, which are used in 
		unit cell reduction and in lattice type determination,
		are studied in \CIII, the space of three complex variables.
		The three complex coordinate planes are composed of the six
		Delone scalars. The transformations at boundaries of the
		Selling reduced orthant are described as matrices of operators.
		A graphical representation as the projections onto the
		three coordinates is described.

@@

	~\citeasnoun{andrews2019b} chose to 
	represent the Selling scalars in the space \SVI{},
	\Svec{} (defined in the order above), 
	as a way to create a metric space
	for the measurement of the distance between lattices. 
	~\citeasnoun{andrews2019b} also considered the representation 
	of this space as the
	space of three complex dimensions, \CIII{} or 
	{$\{c_1$}, {$c_2$}, {$c_3\}$}.

	In \CIII{}, in terms of the Selling scalars, 
	a vector is defined as \{(\si,\siv ), (\sii,\sv),(\siii,\svi)\}, 
	where the real and imaginary parts
	of each are the ``opposite'' scalars 
	according to the definition of~\citeasnoun{Delaunay1932}
	(opposite in terms of the Bravais tetrahedron of scalars) (see~\citeasnoun{andrews2019a}).
	As a mnemonic device, 
	note that the complex components involve ($\alpha,a$), ($\beta, b$), and ($\gamma,c$).
	Additionally, each complex term uses all 
	four 3-space vectors; for example, $c_1$ is ($b \cdot c$, $a \cdot d$).

	~\citeasnoun{andrews2019b} considered the matrix representations 
	of the reflections in \CIII{} (and in \SVI{}). 
	This paper describes the boundary transformations 
	at the edges of the fundamental	unit of \CIII{}
	and also a graphical presentation. 
	In \SVI{}, the fundamental unit is the all negative orthant, 
	which contains only, and all, of the reduced cells. 
	In \SVI{} and \CIII{}, 
	the boundaries located where any \SVI{} scalar 
	(or correspondingly in \CIII{}, a real or imaginary part) 
	equals to zero.

@@

	Next we define the operators in \CIII{} that 
	will be used in the matrix descriptions of 
	the transformations at the boundaries of the fundamental unit,
	see Table ~\ref{tab:operators}.

@@

	\section{Matrices of boundary transformations}
	~~\\

@@

	~\citeasnoun{Delaunay1932} does not 
	consider the boundary transformations 
	in detail. 
	However, he uses them to 
	define the process of Selling reduction. 
	For example in \SVI{}, 
	he lists the following as one of the possible results 
	for a transformation on \si{}, translated to \SVI{}:
	\SvecA{}.
	The third boundary transform for $s_1$ above
	implements this operation and interchanges the
	real part of \ciii{} and the imaginary part of \cii{}:

@@

	Considered in \CIII, Delone's alternate transformation 
	for the \si{} boundary would
	exchange the real of \cii{} 
	with the imaginary part of \ciii{}. That is the fourth
	matrix above in the list for \si{}. 
	The other two transformations for \si{} can 
	be generated from the two we have 
	just mentioned by the ``exchange operation''~\cite{andrews2019b} 
	applied to the second and third \CIII{} coordinates. Delone 
	did not describe the latter two transformations, 
	perhaps because even a single transformation 
	was adequate to implement reduction. He had already listed two.

@@

	The two-dimensional nature of the three coordinates of \CIII{} 
	suggests their use for graphical display.

	As an example, we use Phospholipase A2 (\textbf{PLA2}) (retrieved from 
	the Protein Data Bank \cite{Bernstein1977}), which has had several similar
	or identical structures determined \cite{LeTrong2007}. \citeasnoun{andrews2019b}
	found additional cases (see Table \ref{PLA2})

@@

Below, Figure \ref{fig1} shows the unit cells as 
reported (the centering of lattices has not been removed).
The following figures (Figures \ref{fig2} to \ref{fig5}) show various transformations and
embellishments of the reported cells.

@@

\begin{figure}
	\includegraphics[width=1.0\textwidth]{2_06}
	\caption{The PLA2 unit cells Niggli-reduced. The similarity 
	of the 3 projections is indicative of the exact or nearly
exact rhombohedral symmetry.}
	\label{fig2}
\end{figure}
\begin{figure}
	\includegraphics[width=1.0\textwidth]{3_03}
	\caption{The PLA2 unit cells Delone-reduced. }
	\label{fig3}
\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{4_13}
	\caption{The PLA2 unit cells, Niggli-reduced, and five copies were perturbed 2\%
	orthogonally to \SVI{} vector. }
	\label{fig4}
\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{5_21}
	\caption{The PLA2 unit cells Niggli-reduced, and five copies were perturbed 10\%
		orthogonally to \SVI{} vector. }
	\label{fig5}
\end{figure}

@@

The transformation matrices shown above demonstrate the 
considerable regularity of Selling reduction, as used by 
Delaunay (1932); The reduction operations of Niggli 
reduction \cite{Niggli1928} are more complex. 
While all the boundaries of the non-positive 
orthant of \SVI{} are essentially the same (and 
related by reflections), the boundaries formed in 
Niggli reduction are of multiple types, and the fundamental unit 
of the space representing Niggli-reduced cells (\GVI{}, see \citeasnoun{Andrews2014}) is non-convex.

Several aspects of \CIII{} are evident from inspecting the matrices. 
First, each boundary has four possible 
transformations that can be applied. Since each of 
the transformations at boundaries are self-inverse, they are the 
same transformations that would be used in the process of 
cell (lattice) reduction. \citeasnoun{Delaunay1932} and \citeasnoun{Delone1975} give only two choices, presumably for simplicity, 
omitting transformations that use the ``exchange'' operator 
(see \citeasnoun{andrews2019b}).

@@

	The $C^{++}$ ~code for \CIII{} and related 
	software tools is available in github.com, in
	\url{https://github.com/duck10/LatticeRepLib.git}.

@@

	\bibliography{Reduced}

@@

%C:\Users\lca\Source\Repos\LatticeRepLib\x64\Debug>plotc3
%; Graphical output SVG file =PLT__2023-03-07.13_43_35.svg
%
%C:\Users\lca\Source\Repos\LatticeRepLib\x64\Debug>cmdniggli | plotc3
%; Graphical output SVG file =PLT__2023-03-07.13_44_06.svg
%
%C:\Users\lca\Source\Repos\LatticeRepLib\x64\Debug>cmddelone | plotc3
%; Graphical output SVG file =PLT__2023-03-08.07_11_03.svg
%
%C:\Users\lca\Source\Repos\LatticeRepLib\x64\Debug>cmdniggli | cmdperturb 5 20 | plotc3
%; Graphical output SVG file =PLT__2023-03-08.09_00_13.svg
%
%C:\Users\lca\Source\Repos\LatticeRepLib\x64\Debug>cmdniggli | cmdperturb 5 100 | plotc3
%; Graphical output SVG file =PLT__2023-03-08.09_00_21.svg
